Zohmg---a large scale data store for aggregated time-series-based data. [DRAFT]

### 2009-05-13 doc/report/msc-report.tex
concepts in data warehousing +DONE add running example - aerodyne +DONE concepts in data warehousing: write intro, summary. +DONE concepts: is our use of 'hypercube' standard? => yes, common OLAP term. +DONE concepts: explain slice&dice clearer.

This chapter will serve as an introduction to various concepts in data
warehousing.

@@

\section{Data Warehouse implementations}

@@

At the core of any OLAP system is the OLAP cube (also known as a
multidimensional cube or a hypercube). It consists of numeric facts called
measures which are categorized by dimensions.

The cube is typically modelled in a relational database with fact tables
and dimension tables, upon which the system performs joins to extract data.
\cite{olap_solutions}

@@

A data cube is a conceptual n-dimensional array of values, in which the values
are numerical facts belonging to some dimension. The data cube is also known as
a hypercube, relating to multidimensional analogues of squares and cubes. In
this context, each dimension of the cube corresponds to an attribute of the
data in question and each point along the dimension's axis is a value of the
corresponding attribute. The mental model of a data cube aids in reasoning
about dimensions. Data cubes are traditionally modeled as star schemas in
relational database systems. \cite{olap_solutions}

\subsection{Dimensions}

The school book example of a data cube is three dimensional where each of
the three dimensions correspond to one of the attributes time, products,
and stores. Adding another attribute, for instance sales, adds another
dimension to the data cube, because of the one to one relationship of
attributes and dimensions.

@@

One of the most common operations on a data cube is the projection. A
projection is a dimensionality reduction, where one or more dimensions are
discarded from the cube with the measurements along each axis summed up,
leaving a data cube of a lesser dimension.

A typical projection is that from an n-dimensional cube down to two
dimensions, which is the case when plotting an object---with three or more
dimensions---on screen or paper.

@@

% TODO: example.

@@

\section{Time series}

Time series is one of the most important measures in log analyzing, it
is also very natural when the data is time-stamped. \cite{discoveringweb}

@@

% TODO: be clear about how aggregates are *derived* from one or several vars.
% TODO: example.

An aggregate is a composite of several values. Typical examples of
aggregate functions are average, count, and sum.

@@

>>>>

This chapter introduces data warehouses and the concept of data cubes. It then
introduces common operations on data cubes: projections, slicing and dicing,
and aggregation.

@@

@@

At the core of any OLAP system is the OLAP cube (also known as a
multi-dimensional cube or a hypercube). The cube consists of numeric facts
called measures which are categorized by dimensions. For example, the number
of nails sold is a measure while country, month and type of nail are
dimensions. It is from OLAP that we have appropriated the concept of a data
cube.

@@

A data cube is a conceptual n-dimensional array of values. It aids in
reasoning about dimensions. In the context of this paper, each dimension of
the cube corresponds to an attribute of the data in question and each point
along the dimension's axis is a value of the corresponding attribute. At any
point where the n dimensions intersect, a measure can be found. For example,
if the dimensions are country and nail type, the point in the cube where
country is equal to Sweden and nail type is equal to tack will contain a
measurement of the units of tack nails sold in Sweden.

Data cubes are traditionally modeled as star schemas in relational database
systems. \cite{olap_solutions}

@@

One of the most common operations on a data cube is the projection. A
projection is a function to another data cube, where one or more dimensions
are discarded from the cube and the measurements along each dimensions are
summed up, leaving a data cube of a lesser dimension.

For example, consider the three-dimensional data cube that represents sales of
nails; the dimensions are nail type, country and month. A projection from that
cube to one with two dimensions -- month and nail type -- would flatten the
country dimension: the measurements in the new cube represent the global sales
of each nail type for each month.

@@

@@

For example, consider again the three-dimensional data cube that represents
sales of nails; the dimensions are nail type, country and month. A slice of
that cube with the country dimension fixed to Germany would contains sales
figures for each nail type and month for Germany only.

@@

An aggregate is a composite value that is derived from several variables.
Typical examples of aggregate functions are average, count, and sum.

For example, sales figures for a whole year may be summed up and stored along
the year dimension.

@@

\section{Summary}

This chapter introduced Data Warehouses in general and OLAP in particular. The
concept of data cubes was introduced by OLAP. The chapter was rounded off with
a discussion of dimensionality reducing operations and aggregations of data
cubes.
